---
title: "E-Phac_Scripts"
author: "Athena"
date: "2025-04-10"
output: html_document
editor_options: 
  chunk_output_type: console
---

Useful Libraries & Functions

```{r}
setwd("/Volumes/staff/ephacoffice/DColameo/DATA/human_AG_Sinergia")

source("/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/Adrenal_neededfunctions.R")

```


1. Imaging

A. Load databases

```{r}

getwd()

df_image <- extract_imaging_cols(
  pathDB    = "/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/imaging_data/Output_n1_12.03.db",
  analysis = "pa",
  id_cols = c("Date", "Plate_ID", "Well", "Image_ID", "Channel_Name","Selection", "Selection_Area"),
  num_cols = c("Area", "Mean", "IntDen"),
  scale_num = TRUE,
  scale_cols = c("Area", "Mean", "IntDen")
)

dplyr::glimpse(df_image)

```



B.Image Validation check (4 images from same well)

```{r}

imageval_alt <- function(img.path, idx) {
  
  load_bright <- function(file, factor)
  brighten_image(readImage(file), factor = factor)
  
  # helper function: take one channel and put it in the desired colour slot
  make_grob <- function(img, src_slice = NULL, colour = c("red","green","blue")) {
    if (is.null(src_slice)) {                        # full RGB
      x <- normalize(img)
    } else {                                         # monochrome as chosen colour
      colour  <- match.arg(colour)
      chan    <- normalize(img[,,src_slice])
      rgb_arr <- array(0, dim = c(dim(chan), 3))
      rgb_arr[,, match(colour, c("red","green","blue")) ] <- chan
      x <- rgb_arr
    }
    rasterGrob(x, interpolate = TRUE)
  }
  
  ## --------------------------------------------------------------------
  img.list <- list.files(img.path, pattern = "\\.tif$", recursive = TRUE,
                         full.names = TRUE)
  imgs <- img.list[grepl(idx, img.list)]

  ## ---- bright‑field / BF image (first file) ---------------------------
  img1 <- load_bright(imgs[1], factor = 2)
  bf_channel <- normalize(img1)              
  img1_grob <- make_grob(rotate(img1, 180)) 

  ## ---- fluorescence image (second file) -------------------------------
  img2 <- load_bright(imgs[2], factor = 40)
  img2_grob_green <- make_grob(img2, src_slice = 2, colour = "green")  # plane 2 → green
  img2_grob_red   <- make_grob(img2, src_slice = 3, colour = "red")    # plane 3 → red

  ## composite: R=red, G=green, B=bright‑field
  comp_rgb <- array(0, dim = c(dim(img2[,,3]), 3))
  comp_rgb[,,1] <- normalize(img2[,,3])
  comp_rgb[,,2] <- normalize(img2[,,2])
  comp_rgb[,,3] <- bf_channel
  img2_grob_color <- rasterGrob(comp_rgb, interpolate = TRUE)

  ## ---- arrange --------------------------------------------------------
  grid.arrange(img1_grob, img2_grob_color, img2_grob_green, img2_grob_red, ncol = 2)
}

imageval_alt("/Volumes/staff/ephacoffice/DColameo/DATA/human_AG_Sinergia/hAG_CD56-PE_Cyp11b2-A647_120325_2ndRun_18T39273_HP-Donor/Particle_Analysis/hAG_CD56-PE_Cyp11b2-A647_120325_2ndRun_18T39273_HP-Donor", "H14-1")

```



C. Assigning Cell Color to Well

```{r}

assigncell <- function(PAoutput_path) {
 
  df <- extract_imaging_cols(
  pathDB    = PAoutput_path,
  analysis = "pa",
  id_cols = c("Date", "Plate_ID", "Well", "Image_ID", "Channel_Name","Selection", "Selection_Area"),
  num_cols = c("Area", "Mean", "IntDen"),
  scale_num = TRUE,
  scale_cols = c("Area", "Mean", "IntDen")
)
  
  df$Image_Type <- ifelse(df$Image_ID %% 2 != 0, "fluor", "bf")
  
  df$Channel <- ifelse(df$Channel_Name == "C1", "DAPI", 
                       ifelse(df$Channel_Name == "C2", "Green",
                              ifelse(df$Channel_Name == "C3", "Red", NA)))
  
  df_filtered <- df %>% 
    filter(Image_Type == "fluor", Channel_Name %in% c("C2", "C3"), !is.na(CorrSel), CorrSel == "Hole_ROI")
  
  well_summary <- df_filtered %>%
    group_by(Date, Plate_ID, Image_ID, Well, Channel_Name) %>%
    summarize(Mean = mean(Mean, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = Channel_Name, values_from = Mean, names_prefix = "Mean_")
  
  well_summary <- well_summary %>%
    rename(Mean_Red = Mean_C3,
           Mean_Green = Mean_C2)
  
  well_summary$Mean_Red[is.na(well_summary$Mean_Red)] <- 0
  well_summary$Mean_Green[is.na(well_summary$Mean_Green)] <- 0
  
  well_summary$Row <- substring(well_summary$Well, 1, 1)
  well_summary$Column <- as.numeric(gsub("^[A-Z]+(\\d+)-.*", "\\1", well_summary$Well))
  
  well_summary <- well_summary %>%
    group_by(Column) %>%
    mutate(
      red_range = max(Mean_Red, na.rm = TRUE) - min(Mean_Red, na.rm = TRUE),
      green_range = max(Mean_Green, na.rm = TRUE) - min(Mean_Green, na.rm = TRUE),
      Mean_Red_scaled = ifelse(red_range == 0, 0, (Mean_Red - min(Mean_Red, na.rm = TRUE)) / red_range),
      Mean_Green_scaled = ifelse(green_range == 0, 0, (Mean_Green - min(Mean_Green, na.rm = TRUE)) / green_range)
    ) %>%
    ungroup()
  
  min_active <- 0.05
  
  non_zero_red <- well_summary$Mean_Red_scaled[well_summary$Mean_Red_scaled > 0]
  non_zero_green <- well_summary$Mean_Green_scaled[well_summary$Mean_Green_scaled > 0]
  
  red_threshold <- if (length(non_zero_red) > 0) {
    quantile(non_zero_red, 0.4, na.rm = TRUE)
  } else {
    0 
  }
  
  green_threshold <- 0.1
  
  well_summary <- well_summary %>%
    mutate(Color = case_when(
      (Mean_Red_scaled < min_active & Mean_Green_scaled < min_active) ~ "black",
      (Mean_Red_scaled >= red_threshold & Mean_Green_scaled >= green_threshold) ~ "yellow",
      (Mean_Green_scaled >= green_threshold) ~ "green",
      (Mean_Red_scaled >= red_threshold) ~ "red",
      TRUE ~ "black"
    )) 
  
  return (well_summary)
  
}

directory <- "/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/imaging_data"
extension <- "\\.db$"  # Regular expression for files ending with .db

file_list <- list.files(path = directory, pattern = extension, full.names = TRUE)

well_df <- tibble()

for (file in file_list) {
  ws <- assigncell(file)
  ws$file_name <- file
  well_df <- bind_rows(well_df, ws)
  
  p <- ggplot(ws, aes(x = Column, y = Row, fill = Color)) +
  geom_tile(color = "black", linewidth = 0.5) +
  scale_fill_manual(values = c("red" = "red", "green" = "green", "yellow" = "yellow", "black" = "grey")) +
  scale_x_continuous(breaks = 1:24) +
  scale_y_discrete(limits = rev(sort(unique(well_df$Row)))) +
  geom_text(aes(label = Well), color = "black") +
  labs(title = paste("Well Plate Cell Assignment for", ws$Date, ws$Plate_ID),
  x = "Column", y = "Row")
  
  print(p)
    
}


```


Merging of the Imaging and Electrophysiological Data

A. Data Prep

```{r}

# wd_path: path to the excel files
wd_path <- "/Volumes/staff/ephacoffice/sharing/Skripts/R-Skripts/Human_adrenal_glands"

l_files <- list.files(path = wd_path ,pattern = "*.xlsx$", recursive = TRUE, full.names = TRUE)

l_files <- l_files[!grepl("~", l_files)]
meta <- l_files[grepl("Experiment", l_files)]
l_files <- l_files[!grepl("Experiment", l_files)]
length(l_files)

dfs <- lapply(l_files, function(x) {
  df <- prepareDF_comb(x)
  
  return(df)
})

names(dfs) <- l_files
df <- bind_rows(dfs, .id = "column_label")
head(df)
df$Plate_ID <- sapply(df$Plate_ID, function(x){
  unlist(str_split(x, "\\r"))[1]
})

meta <- as.data.frame(read_excel(meta))
unique(meta$Plate_ID)

df <- merge(df, meta, by="Plate_ID")
unique(df$Plate_ID)

df$Row <- sapply(df$Well, function(x){
  str_sub(x, 1, 1)
})

df$Column <- sapply(df$Well, function(x){
  as.numeric(str_sub(x, 2, 3))
})

letter2number <- function(x) {utf8ToInt(x) - utf8ToInt("A") + 1L}
df$RowNum <- sapply(df$Row, function(c){letter2number(c)})

df$Na_Conc <- ifelse(df$Sweep <=24, 0.08, 
                     ifelse(df$Sweep >24 & df$Sweep <= 42,  0.04, 
                     ifelse(df$Sweep >42 & df$Sweep <= 60,  0.02, 
                     ifelse(df$Sweep >60 & df$Sweep <= 84, 0.01, 0))))

df$Compound <- ifelse(df$RowNum %in% 1:4 | df$RowNum %in% 9:12, "Na Addition", "Na Removal")

df$Na_Conc <- ifelse(df$Sweep <=24 & df$Compound == "Na Addition", 0.08, 
                     ifelse(df$Sweep >24 & df$Sweep <= 42 & df$Compound == "Na Addition",  0.1, 
                     ifelse(df$Sweep >42 & df$Sweep <= 60& df$Compound == "Na Addition",  0.12, 
                     ifelse(df$Sweep >60 & df$Sweep <= 84& df$Compound == "Na Addition", 0.13, df$Na_Conc))))

unique(df$Conditions)

df$Group <- ifelse(as.numeric(df$Column)%%2 == 0, "WT", "HP")
df$Group <- ifelse(df$Conditions == "Donor" | df$Conditions == "Donor1|Donor2", "WT", df$Group)

unique(df$Group)

unique(df$Plate_ID)

# df <- df %>% filter(Rec_Date == "12.03.2025")

saveRDS(df, "/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/combined_df.rds")

```



```{r}

feature_cols <- c("X.100mV", "X.20mV", "Erev", "pA.pF", "Cap", "Rseal")

long_df <- df %>%
  pivot_longer(
    cols      = all_of(feature_cols),
    names_to  = "Feature",          # e.g. "X.100mV"
    values_to = "Value"
  )

main_tables <- long_df %>%
  unite(col = "SweepFeature", Sweep, Feature, sep = "_") %>%   # "1_X.100mV"
  pivot_wider(
    id_cols      = c(Plate_ID, Well),            # keep Wells distinct
    names_from   = SweepFeature,
    values_from  = Value
  ) %>%
  group_split(Plate_ID) %>%              # a list of data frames
  setNames(unique(long_df$Plate_ID))     # you can access with main_tables[["PlateA"]]

nested_tbl <- long_df %>%
  group_by(Plate_ID, Feature) %>%        # one row == one (Plate, Feature)
  nest() %>%                             # 'data' list-column holds Well/Sweep/Value
  mutate(
    # inside each little tibble: make Sweeps columns so it is Well × Sweep
    data = map(data, ~ pivot_wider(.x,
                                   id_cols     = Well,
                                   names_from  = Sweep,
                                   values_from = Value,
                                   names_repair = "minimal"
                                   ))
  )

```



B. Per cell:
```{r}
df <- readRDS("/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/combined_df.rds")

df <- ag(subset(df), cols= c("Well", "QC", "Compound", "Group", "Na_Conc", "Plate_ID", "Group"))

df$Erev <- 1000*df$Erev
df$X.100mV <- 10^12*df$X.100mV

metric <- "Erev"

df_wide <- dcast(setDT(df), Well + QC + Group + Compound + Plate_ID ~ Na_Conc, value.var = metric)

df_wide[df_wide$QC == TRUE,]
names(df_wide)[6:12] <- paste("NaC", 1000*as.numeric(names(df_wide)[6:12]), sep=".")

df_wide$QC <- ifelse(df_wide$NaC.80 <= -30 & df_wide$QC == TRUE, TRUE, FALSE)
df_wide[df_wide$QC == TRUE,]
df_wide <- as.data.frame(df_wide)

df_wide$Add1 <- NA
df_wide$Add2 <-  NA
df_wide$Add3 <-  NA
df_wide$Add4 <- NA

names(df_wide)
df_wide

for(row in 1:nrow(df_wide)){
  print(row)
  
  if(df_wide[row,]$Compound == "Na Removal"){
    df_wide[row,13:16] <- c(df_wide[row,"NaC.80"],df_wide[row,"NaC.40"],df_wide[row,"NaC.20"], df_wide[row,"NaC.10"] )
    
  } else{
    df_wide[row,13:16] <- df_wide[row,9:12]
    
  }

}

View(df_wide)

df_wide$Row <- substring(df_wide$Well, 1, 1)
df_wide$Column <- substring(df_wide$Well, 2, 3)

df_melt <- reshape2::melt(df_wide, measure.var = c("Add1", "Add2", "Add3", "Add4"), 
                              id.var = c("Well", "QC", "Compound", "Group", "Plate_ID"))

df_melt$Addition <- sapply(df_melt$variable, function(x){
  as.numeric(str_sub(x, 4, 4))
})
```


```{r}

## Clustering

df_wide <- subset(df_wide, QC == TRUE)

df_scaled <- scale(df_wide[, grepl("Add", colnames(df_wide))])


rownames(df_scaled) <- paste(df_wide$Well, df_wide$Plate_ID, sep=".")

df_scaled <- df_scaled[complete.cases(df_scaled),]

df.umap <- umap(df_scaled)
head(df.umap$layout, 3)

umap_df <- df.umap$layout %>%
  as.data.frame()%>%
  rename(UMAP1="V1",
         UMAP2="V2") 

k <- 3
kc <- kmeans(umap_df[,c("UMAP1", "UMAP2")], k)

umap_df$Well <- rownames(umap_df)
ggplot(umap_df, aes(x = UMAP1, 
             y = UMAP2,color=as.factor( kc$cluster)))+
  geom_point(size=2) + theme_minimal(base_size = 18) +labs(color = "Clusters") 

umap_df$Plate <- sapply(umap_df$Well, function(x){
  unlist(str_split(x, "\\."))[2]
})

ggplot(umap_df, aes(x = UMAP1, 
             y = UMAP2,color=as.factor(Plate)))+
  geom_point(size=2) + theme_minimal(base_size = 18) +labs(color = "Plate ID") 

clusters <- as.data.frame(factor(kc$cluster))
clusters$Well.Plate <- rownames(clusters)
clusters$cluster <- clusters$`factor(kc$cluster)`
df_melt$Well.Plate <- paste(df_melt$Well, df_melt$Plate_ID, sep=".")

df_melt <- merge(df_melt, clusters, by="Well.Plate")

nsum <- subset(df_melt, QC == TRUE & Compound == "Na Removal" & Addition == 1) %>%
  group_by(QC, Group, Compound, cluster) %>%
  summarise(n = n())

nsum

```



```{r}
# Clustering based on Add1

valid_idx   <- which(!is.na(df_wide$Add1))
add1_vec  <- df_wide$Add1[valid_idx]
add1_vec <- scale(add1_vec)
set.seed(42)

um <- umap(matrix(add1_vec, ncol = 1),
           n_neighbors = 15,
           min_dist    = 0.1,
           metric      = "euclidean")

umap_df <- data.frame(UMAP1 = um$layout[,1],
                      UMAP2 = um$layout[,2])

km_add1  <- kmeans(umap_df, centers = 3, nstart = 50)   # adjust centres

df_wide$cluster_min <- NA_integer_
df_wide$cluster_min[!is.na(df_wide$Add1)] <- km_add1$cluster
umap_df$cluster <- factor(km_add1$cluster)

cluster_means <- tapply(add1_vec, km_add1$cluster, mean)
  
cluster_plot <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = cluster)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_brewer(palette = "Set1", name = "Clusters") +
  labs(
    title    = "UMAP of Add1 only Clustering",
    subtitle = "Each point = one cell",
    x        = "UMAP1", y = "UMAP2"
  ) +
  theme_minimal()
  
print(cluster_plot)

dist_mat <- dist(add1_vec)
sil      <- silhouette(km_add1$cluster, dist_mat)
cat("Mean silhouette width:", round(mean(sil[, "sil_width"]), 3), "\n")
```


```{r}
# Clustering based on Add2

valid_idx   <- which(!is.na(df_wide$Add2))
add2_vec  <- df_wide$Add2[valid_idx]
add2_vec  <- scale(add2_vec)
set.seed(42)

um <- umap(matrix(add2_vec, ncol = 1),
           n_neighbors = 15,
           min_dist    = 0.1,
           metric      = "euclidean")

umap_df <- data.frame(UMAP1 = um$layout[,1],
                      UMAP2 = um$layout[,2])

km_add2  <- kmeans(umap_df, centers = 3, nstart = 50)   # adjust centres

df_wide$cluster_min <- NA_integer_
df_wide$cluster_min[!is.na(df_wide$Add2)] <- km_add2$cluster
umap_df$cluster <- factor(km_add2$cluster)

cluster_means <- tapply(add2_vec, km_add2$cluster, mean)

cluster_plot <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = cluster)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_brewer(palette = "Set1", name = "Clusters") +
  labs(
    title    = "UMAP of Add2 only Clustering",
    subtitle = "Each point = one cell",
    x        = "UMAP1", y = "UMAP2"
  ) +
  theme_minimal()
  
print(cluster_plot)

dist_mat <- dist(add2_vec)
sil      <- silhouette(km_add2$cluster, dist_mat)
cat("Mean silhouette width:", round(mean(sil[, "sil_width"]), 3), "\n")


```


```{r}
# Clustering based on Add3

valid_idx   <- which(!is.na(df_wide$Add3))
add3_vec  <- df_wide$Add3[valid_idx]
add3_vec  <- scale(add3_vec)
set.seed(42)

um <- umap(matrix(add3_vec, ncol = 1),
           n_neighbors = 15,
           min_dist    = 0.1,
           metric      = "euclidean")

umap_df <- data.frame(UMAP1 = um$layout[,1],
                      UMAP2 = um$layout[,2])

km_add3  <- kmeans(umap_df, centers = 3, nstart = 50)   # adjust centres

df_wide$cluster_min <- NA_integer_
df_wide$cluster_min[!is.na(df_wide$Add3)] <- km_add3$cluster
umap_df$cluster <- factor(km_add3$cluster)

cluster_means <- tapply(add3_vec, km_add3$cluster, mean)
  
cluster_plot <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = cluster)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_brewer(palette = "Set1", name = "Clusters") +
  labs(
    title    = "UMAP of Add3 only Clustering",
    subtitle = "Each point = one cell, colored by 1D kmeans on Add3",
    x        = "UMAP1", y = "UMAP2"
  ) +
  theme_minimal()
  
print(cluster_plot)

dist_mat <- dist(add3_vec)
sil      <- silhouette(km_add3$cluster, dist_mat)
cat("Mean silhouette width:", round(mean(sil[, "sil_width"]), 3), "\n")

```

```{r}
# Clustering based on Add4

valid_idx   <- which(!is.na(df_wide$Add4))
add4_vec  <- df_wide$Add4[valid_idx]
add4_vec  <- scale(add4_vec)
set.seed(42)

um <- umap(matrix(add4_vec, ncol = 1),
           n_neighbors = 15,
           min_dist    = 0.1,
           metric      = "euclidean")

umap_df <- data.frame(UMAP1 = um$layout[,1],
                      UMAP2 = um$layout[,2])

km_add4  <- kmeans(umap_df, centers = 3, nstart = 50)   # adjust centres

df_wide$cluster_min <- NA_integer_
df_wide$cluster_min[!is.na(df_wide$Add4)] <- km_add4$cluster
umap_df$cluster <- factor(km_add4$cluster)

cluster_means <- tapply(add4_vec, km_add4$cluster, mean)
  
cluster_plot <- ggplot(umap_df, aes(x = UMAP1, y = UMAP2, color = cluster)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_brewer(palette = "Set1", name = "Clusters") +
  labs(
    title    = "UMAP of Add4 only Clustering",
    subtitle = "Each point = one cell, colored by 1D kmeans on Add4",
    x        = "UMAP1", y = "UMAP2"
  ) +
  theme_minimal()
  
print(cluster_plot)

dist_mat <- dist(add4_vec)
sil      <- silhouette(km_add4$cluster, dist_mat)
cat("Mean silhouette width:", round(mean(sil[, "sil_width"]), 3), "\n")

```


```{r}

# Clustering based on NaC.80, Add1, Add2, Add3, Add4

var_set <- c("Add1", "Add2", "Add3", "Add4")
idx <- complete.cases(df_wide[, var_set])

feat_mat <- df_wide[idx, var_set] %>%          
  scale() %>%                                  
  as.matrix()

set.seed(42)

um <- umap(feat_mat,
           n_neighbors = 15,
           min_dist    = 0.1,
           metric      = "euclidean")

umap_df <- data.frame(UMAP1 = um$layout[,1],
                      UMAP2 = um$layout[,2])

km_multi <- kmeans(umap_df, centers = 3, nstart = 50)   # adjust centres

df_wide$cluster_multi <- NA_integer_
df_wide$cluster_multi[idx] <- km_multi$cluster
umap_df$cluster <- factor(km_multi$cluster)

cluster_plot <- ggplot(umap_df, aes(UMAP1, UMAP2, color = cluster)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_brewer(palette = "Set1", name = "Clusters") +
  labs(
    title    = "UMAP of Na Additions 1-4",
    subtitle = "Each point = one cell",
    x        = "UMAP1", y = "UMAP2"
  ) +
  theme_minimal()

print(cluster_plot)

dist_mat <- dist(feat_mat)
sil      <- silhouette(km_multi$cluster, dist_mat)
cat("Mean silhouette width:", round(mean(sil[, "sil_width"]), 3), "\n")

```



Unsupervised ML

```{r}

library(tidyverse)
library(randomForest)
library(factoextra)
library(umap)

# -------- data prep
X <- df_wide %>%
  select(NaC.80, Add1, Add2, Add3, Add4) %>%
  drop_na() %>%
  scale() %>%
  as.matrix()

# -------- clustering
set.seed(42)
k_best <- 3                               
km      <- kmeans(X, centers = k_best, nstart = 50)
y <- factor(km$cluster)  

# -------- feature importance (URF)
rf_unsup <- rfsrc(                     
  data        = as.data.frame(X),      # X = your scaled feature matrix
  ntree       = 500,
  proximity   = TRUE,                  # or "inbag"/"all"        
  unsupervised = TRUE                  # explicit but not strictly needed
)

md <- max.subtree(rf_unsup, max.order = 1)          # first‑order depths
depth_rank <- sort(md$order[, 1])                  # smaller ⇒ more predictive
print(depth_rank)       # the shallower (smaller) the first‑order depth, the earlier that variable tends to split the trees → the stronger its structural signal in the data hierarchy.


md_vals <- c(Add3 = 1.368, Add2 = 1.502, Add1 = 2.044, NaC.80 = 2.118)

# convert: lower depth → larger score
rel_imp <- max(md_vals) - md_vals
rel_imp <- rel_imp / sum(rel_imp)          # normalise to 0‑1

print(round(rel_imp, 3))


# predict your k‑means class labels; importance answers “what makes cluster A look different from cluster B?”
# Maxima > Minima ≫ Capacitance → Maxima distinguishes clusters best 
km_labels <- factor(km$cluster)      # k‑means clusters as target
rf_sup <- rfsrc(km_labels ~ ., data = data.frame(km_labels, X),
                ntree = 500, importance = "permute")
sort(rf_sup$importance[, "all"], decreasing = TRUE)


```




C. Dataset Merging:

```{r}

df_wide$Column <- as.numeric(df_wide$Column)

well_df$Column <- as.numeric(well_df$Column)

combined_df <- inner_join(df_wide, well_df, by = c("Plate_ID", "Row", "Column"))%>%
  select(-red_range, -green_range, -Well.x, -Well.y)

```


Well Plots by Plate ID (Dual Plot)

```{r}

p1 <- ggplot(subset(combined_df), aes(x=as.numeric(Column), y=Row, fill=Add1)) + geom_tile()+ 
  scale_x_continuous(breaks = 1:24) + 
  scale_y_discrete(limits = rev) + 
  labs(x = "Column") +
  geom_text(aes(label=interaction(Column, Row)), color="black") + facet_grid(~Plate_ID)

p2 <- ggplot(subset(combined_df), aes(x=as.numeric(Column), y=Row, fill=Color)) + geom_tile(color = "black", linewidth = 0.5)+ 
  scale_fill_manual(values = c("red" = "red", "green" = "green", "yellow" = "yellow", "black" = "grey")) +
  scale_x_continuous(breaks = 1:24) +
  scale_y_discrete(limits = rev) +
  labs(x = "Column") +
  geom_text(aes(label=interaction(Column, Row)), color="black") + facet_grid(~Plate_ID) 

grid.arrange(p1, p2, nrow=2)

```


Clustering and significance of the merged data

```{r}

df_cluster <- combined_df %>%
  select(Plate_ID, Add1, Mean_Red_scaled, Mean_Green_scaled, Color, Group) %>%
  na.omit()

df_cluster_scaled <- scale(df_cluster %>% select(-Color, -Plate_ID, -Group))

k <- 4
set.seed(123)  # for reproducibility
kc <- kmeans(df_cluster_scaled, k)

df_cluster$cluster <- as.factor(kc$cluster)

table(Original = df_cluster$Color, Cluster = df_cluster$cluster)

# Clustering

ggplot(df_cluster, aes(x = Add1, y = Mean_Red_scaled, fill = Group, color = Color, shape = cluster)) +
  geom_point(size = 3, stroke = 1) +
  scale_shape_manual(values = c("1" = 21, "2" = 22, "3" = 23, "4" = 24)) +
  scale_fill_manual(values = c("WT" = "light blue", "HP" = "violet"), name   = "Group") +
  guides(
    fill   = guide_legend(override.aes = list(shape = 21,
                                              colour = "black",   # constant outline so fill legend is clear
                                              size = 4)),
    colour = guide_legend(override.aes = list(shape = 21,
                                              fill  = "white",   # hollow keys so outline colours stand out
                                              size  = 4))) +
  facet_wrap(~Plate_ID) +
  labs(title = "K-means Clustering (k = 4)",
       subtitle = "Outline = cluster, fill = HP vs WT, shape = cluster",
       x = "Erev",
       y = "Mean Red Scaled",
       fill  = "Group") +
  scale_color_manual(values = c("red" = "red",
                                "black" = "grey",
                                "yellow" = "yellow",
                                "green" = "green"), 
                    name = "Cluster") +
  theme_minimal()


# Significance calculation

table_color_cluster <- table(df_cluster$Color, df_cluster$cluster)
print(table_color_cluster)

# Fisher's test
fisher_result <- fisher.test(table_color_cluster)
print(fisher_result)


saveRDS(df_cluster, "/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/df_cluster.rds")

```


Function that takes the Plate ID, Row and Column in and returns the assigned color

```{r}

assigned_color <- function(plate_id, row, column, data = combined_df){
  result <- data %>%
    filter(Plate_ID == plate_id, Row == row, Column == column)
  
  if (nrow(result)==0) {
    stop("No matching record found")
  }
  
  return(result$Color)
}

```



