---
title: "ROMK code"
author: "Athena Schumacher"
date: "2025-07-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

ROMK
```{r}
devtools::load_all()

parent_folder <- "/Volumes/staff/ephacoffice/DColameo/DATA/ROMK"
data("se_romk")

wells <- se_romk$Well
plateIDs <- se_romk$Plate_ID

image_paths_list <- mapply(function(well, plate, location) {
  image_paths(parent_folder, idx = well, plate = plate, location = 5)
}, well = wells, plate = plateIDs, SIMPLIFY = FALSE)

colData(se_romk)$Image_paths <- image_paths_list 

# Check 
colData(se_romk)$Image_paths
se_romk[,se_romk$Well == "H05"]$Image_paths
```


```{r}
plot_intensity_distribution(se_romk, channels = c("C1", "C2", "C3"), plot_type="density")
plot_intensity_distribution(se_romk, channels = c("C1", "C2", "C3"), plot_type="ecdf")

se_ROMK_CA <- assign_cell_FINAL(
  SE = se_romk,
  channels = c("C1", "C2", "C3"),
  value_col = "Mean",
  classification_method = "clustering",
  color_output = "Color_cluster",
  plot_umap = TRUE,
  k_centers = 3
)

se_ROMK_CA <- assign_cell_FINAL(
  SE = se_ROMK_CA,
  channels = c("C1", "C2", "C3"),
  chan2analyze = c("C1", "C2", "C3"),
  value_col = "Mean",
  classification_method = "manual",
  color_output = "Color_manual",
  threshold_mode = "manual",
  red_thresh = 7.6,
  green_thresh = 4.6,
  c3_thresh = 6.6
  #red_quant = 0.665,
  #green_quant = 0.665,
  #c3_quant = 0.94
)

plot_well_assignment(se_ROMK_CA)

well <- "K04"

imageval(se_ROMK_CA, well, "22W14457", green_slice = 3, red_slice = 2)

colData(se_ROMK_CA[,se_ROMK_CA$Well == well])
assays(se_ROMK_CA[,se_ROMK_CA$Well == well])$Erev*1000

colData(se_ROMK_CA)[colData(se_ROMK_CA)$Well == well,]

colData(se_ROMK_CA)[colData(se_ROMK_CA)$Well == well, "C3"]


unique(se_ROMK_CA$C3$CorrSel)

```


Filter just for the first addition (optional)
```{r}
colData(se_ROMK_CA)[30:40,]

rowData(se_ROMK_CA)$Additions <- ifelse(rowData(se_ROMK_CA)$Sweep <=24, "Add1",
                     ifelse(rowData(se_ROMK_CA)$Sweep >24 & rowData(se_ROMK_CA)$Sweep <= 42,  "Add2",
                     ifelse(rowData(se_ROMK_CA)$Sweep >42 & rowData(se_ROMK_CA)$Sweep <= 60, "Add3",
                     ifelse(rowData(se_ROMK_CA)$Sweep >60 & rowData(se_ROMK_CA)$Sweep <= 84,"Add4", NA))))

sweeps_to_use <- which(rowData(se_ROMK_CA)$Additions == "Add1")

se_ROMK_CA <- colAG(se_ROMK_CA, assayList = c("Erev"), sweeps = sweeps_to_use)

se_ROMK_CA <- se_ROMK_CA[, complete.cases(se_ROMK_CA$Erev_mean)]

se_ROMK_CA <- se_ROMK_CA[, colData(se_ROMK_CA)$Erev_mean < -0.03 ]

se_ROMK_CA$Erev_mean <- se_ROMK_CA$Erev_mean*1000
```

Include the group assignments in your SummarizedExperiment
```{r}
groups <- c(
  rep("mScarlett", 4),
  rep("GFP", 4),
  rep("Mix", 12),
  rep("Empty", 4)
)

se_ROMK_CA <- group_assignment(se_ROMK_CA, se = "Yes", pattern = "Cycle", cycle_pattern = groups)
```

Clustering solely based on ROMK channel (C3)
```{r}
se_ROMK_assigned_ROMKonly <- assign_cell_FINAL(
  SE = se_romk,
  channels = c("C3"),
  value_col = "Mean",
  classification_method = "clustering",
  color_output = "Color_cluster",
  plot_umap = FALSE,
  k_centers = 4
)
df <- as.data.frame(colData(se_ROMK_assigned_ROMKonly))
ggplot(df, aes(x = cluster, y = feature_C3_scaled, color = cluster)) +
  geom_jitter(width = 0.2, alpha = 0.7, size = 2) +
  theme_minimal() +
  labs(title = "Clustering on C3 only", y = "log1p (C3 Mean)", x = "Cluster")

saveRDS(se_ROMK_assigned_ROMKonly, "/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/SE_ROMKchanonly.rds")
```


Compare the manual and clustering color assignments
```{r}
p1 <- ggplot(df, aes(x = UMAP1, y = UMAP2, color = Color_manual)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_manual(values = c("red" = "red", "green" = "green", "black" = "black")) +
  theme_minimal() +
  labs(
    title = "UMAP of Scaled Features",
    subtitle = "Manual Cell Assignment",
    x = "UMAP1", y = "UMAP2"
  )

p2 <- ggplot(df, aes(x = UMAP1, y = UMAP2, color = Color_cluster)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_brewer(palette = "Set1", name = "Cluster") +
  theme_minimal() +
  labs(
    title = "UMAP of Scaled Features",
    subtitle = "Clusters assigned via k-means on UMAP",
    x = "UMAP1", y = "UMAP2"
  )

grid.arrange(p1, p2, ncol=2)
```


Go through the individual channels and look at the overall induction distribution
```{r}
col_df <- as.data.frame(colData(se_ROMK_CA))

thresh_list <- c(C1 = 0.67, C2 = 0.65, C3 = 0.48)

for (channel in names(thresh_list)) {
  feat_col <- paste0("feature_", channel, "_scaled")
  bin_col <- paste0(channel, "_binary")
  
  if (!feat_col %in% colnames(col_df)) {
    warning(paste("Column", feat_col, "not found in colData. Skipping."))
    next
  }
  col_df[[bin_col]] <- ifelse(col_df[[feat_col]] > thresh_list[[channel]], 1, 0)
}

for (col in setdiff(colnames(col_df), colnames(colData(se_ROMK_CA)))) {
  colData(se_ROMK_CA)[[col]] <- col_df[[col]]
}

colData(se_ROMK_CA)$Overall_Positive <- with(col_df, ifelse(C1_binary == 1 & C2_binary == 1 & C3_binary == 1, 1, 0))

col_df$Color_binary <- with(col_df, case_when(
  C1_binary == 1 & C2_binary == 1 & C3_binary == 1 ~ "white", #not-induced
  C1_binary == 1 & C2_binary == 1 & C3_binary == 0 ~ "yellow", #ambiguous
  C1_binary == 1 & C2_binary == 0 & C3_binary == 1 ~ "turquoise", #induced
  C1_binary == 0 & C2_binary == 1 & C3_binary == 1 ~ "orange", #ambiguous
  C1_binary == 1 & C2_binary == 0 & C3_binary == 0 ~ "green", #induced
  C1_binary == 0 & C2_binary == 1 & C3_binary == 0 ~ "red", #not induced
  C1_binary == 0 & C2_binary == 0 & C3_binary == 1 ~ "navy", #induced
  TRUE ~ "black" 
))

colData(se_ROMK_CA)$Color_binary <- col_df$Color_binary

col_df$Induction_status <- with(col_df, case_when(
  Color_binary %in% c("turquoise", "green", "navy") ~ "induced",
  Color_binary %in% c("white", "red") ~ "not_induced",
  Color_binary == c("yellow", "orange") ~ "ambiguous",
  TRUE ~ "unknown"
))

col_df$Induction_status <- factor(col_df$Induction_status,
                                   levels = c("induced", "not_induced", "ambiguous", "unknown"))

colData(se_ROMK_CA)$Induction_status <- col_df$Induction_status

plot_well_binary_assignment(se_ROMK_CA, values = c(
        "red" = "red",
        "green" = "green",
        "navy" = "navy",
        "yellow" = "yellow",
        "orange" = "orange",
        "turquoise" = "turquoise",
        "white" = "white",
        "black" = "grey"
      ))

plot_well_binary_assignment(se_ROMK_CA, color_col = "Induction_status", values = c(
        "induced" = "steelblue",
        "not_induced" = "tomato",
        "ambiguous" = "gold",
        "unknown" = "grey"
      ) , title = "Induction Status Map")

colData(se_ROMK_CA)[40:45, 158:162]
```



Compare ROMK expression
ROMK expression in mScarlet+ cells (1:4)
```{r}

col_df <- as.data.frame(colData(se_ROMK_CA))
col_df$Column <- as.numeric(col_df$Column)

# Subset to wells in columns 21 to 24
subset_mScar <- col_df[col_df$Column %in% 1:4, ]

mean_C3_mScar <- mean(subset_mScar$feature_C3_scaled, na.rm = TRUE)
mean_C3_mScar
#2.911363

median_C3_mScar <- median(subset_mScar$feature_C3_scaled, na.rm = TRUE)
median_C3_mScar
#4.805203

```

ROMK expression in GFP+ cells (5:8)
```{r}

# Subset to wells in columns 5 to 8
subset_GFP <- col_df[col_df$Column %in% 5:8, ]

mean_C3_cols_GFP <- mean(subset_GFP$feature_C3_scaled, na.rm = TRUE)
mean_C3_cols_GFP
#3.901818

median_C3_cols_GFP <- median(subset_GFP$feature_C3_scaled, na.rm = TRUE)
median_C3_cols_GFP
#5.799439

```

ROMK expression in empty cells (21:24)
```{r}

# Subset to wells in columns 21 to 24
subset_df <- col_df[col_df$Column %in% 21:24, ]

mean_C3_last_cols <- mean(subset_df$feature_C3_scaled, na.rm = TRUE)
mean_C3_last_cols
#3.484095

median_C3_last_cols <- median(subset_df$feature_C3_scaled, na.rm = TRUE)
median_C3_last_cols
#5.229206

```

```{r}
subset_mScar$Group <- "mScar (Cols 1–4)"
subset_GFP$Group <- "GFP (Cols 5–8)"
subset_df$Group <- "Last Cols (21–24)"

# Combine all subsets into one dataframe
combined_df <- rbind(subset_mScar, subset_GFP, subset_df)

ggplot(combined_df, aes(x = Group, y = feature_C3_scaled, fill = Group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black") +
  theme_minimal() +
  labs(title = "Comparison of Scaled Feature C3 Across Groups",
       y = "Scaled C3", x = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "none") +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 3, color = "red", fill = "red") +
  scale_fill_manual(values = c("mScar (Cols 1–4)" = "coral", "GFP (Cols 5–8)" = "forestgreen", "Last Cols (21–24)" = "lightblue"
  )) 
```


```{r}
col_df$Column_Group <- case_when(
  col_df$Column %in% 1:4 ~ "Cols_1_4",
  col_df$Column %in% 5:8  ~ "Cols_5_8",
  col_df$Column %in% 21:24 ~ "Cols_21_24",
  TRUE ~ NA_character_
)

# Filter only those two groups
plot_df <- subset(col_df, !is.na(Column_Group))

# Boxplot
ggplot(plot_df, aes(x = Column_Group, y = feature_C3_scaled, fill = Column_Group)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Comparison of feature_C3_scaled", y = "C3 Scaled", x = "Column Group")

#################

col_df <- as.data.frame(colData(se_ROMK_assigned))
col_df$Column <- as.numeric(col_df$Column)

# Add grouping variable
col_df <- col_df %>%
  mutate(Column_Group = case_when(
    Column %in% 5:8 ~ "Cols_5_8",
    Column %in% 21:24 ~ "Cols_21_24",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(Column_Group))  # Keep only wells from those column groups

# Pivot to long format for the 3 channels
long_df <- col_df %>%
  pivot_longer(
    cols = c(feature_C1_scaled, feature_C2_scaled, feature_C3_scaled),
    names_to = "Channel",
    values_to = "Scaled_Intensity"
  ) %>%
  mutate(Channel = gsub("feature_|_scaled", "", Channel))  # clean channel labels


ggplot(long_df, aes(x = Column_Group, y = Scaled_Intensity, fill = Column_Group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.8) +
  geom_jitter(width = 0.1, alpha = 0.4, size = 1) +
  facet_grid(. ~ Channel) +
  theme_minimal() +
  labs(
    title = "Comparison of Scaled Intensities Across Channels",
    x = "Column Group",
    y = "Scaled Intensity"
  ) +
  theme(legend.position = "none")

```



Looking at all the cells
```{r}
path <- "/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/ROMK/NCI_ramp_ATP1A1_18.30.46.xlsx"

# romk_excel <- prepareDF("/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/ROMK/NCI_ramp_ATP1A1_18.30.46.xlsx")
# romk_excel$Erev <- as.numeric(romk_excel$Erev)
# romk_excel$Erev <- romk_excel$Erev*1000
# romk_se <- prepareSE(path)

l_files <- list.files(path = "/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/ROMK" ,pattern = "*.xlsx$", recursive = TRUE, full.names = TRUE)
romk_se <- prepareSE(l_files)

unique(romk_se$C3.C2$CorrSel)

romk <- romk_se
wells <- romk$Well
plateIDs <- romk$Plate_ID

image_paths_list <- mapply(function(well, plate, location) {
  image_paths(parent_folder, idx = well, plate = plate, location = 5)
}, well = wells, plate = plateIDs, SIMPLIFY = FALSE)

colData(romk)$Image_paths <- image_paths_list 
# Check 
colData(romk)$Image_paths
romk[,romk$Well == "H05"]$Image_paths

plot_intensity_distribution(romk_col, channels = c("C1.C3.hole", "C2.C3.hole"), plot_type="density", channel_colors = c(C1.C3.hole = "forestgreen", C2.C3.hole = "lightcoral"))
plot_intensity_distribution(romk_col, channels = c("C1.C3.hole", "C2.C3.hole"), plot_type="ecdf", channel_colors = c(C1.C3.hole = "forestgreen", C2.C3.hole = "lightcoral"))
plot_intensity_distribution(romk_col, channels = "C2.C3.background", plot_type="ecdf")

ROMK_pa_c3 <- assign_cell_FINAL_valcol(
  SE = romk_pa,
  channels = c("C2.background", "C3.hole", "C2.hole"),
  value_col = c("Mean", "Area"),
  classification_method = "clustering",
  color_output = "Color_cluster",
  plot_umap = TRUE,
  k_centers = 3
)

ROMK_pa_c3 <- assign_cell_FINAL(
  SE = ROMK_pa_c3,
  channels = c("C3.hole", "C3.background"),
  chan2analyze = c("C3.hole", "C3.background"),
  value_col = "Mean",
  classification_method = "manual",
  color_output = "Color_manual",
  threshold_mode = "manual",
  green_thresh = 5.5,
  red_thresh = 5.8,
  c3_thresh = 6.6
  #red_quant = 0.665,
  #green_quant = 0.665,
  #c3_quant = 0.94
)

plot_well_assignment(ROMK_pa_c3, color_col = "Color_cluster")
plot_well_assignment(ROMK_pa_c3, color_col = "Color_cluster", colors = c("1" = "grey", "3" = "red", "2" = "green"))

saveRDS(ROMK_pa_c3, "/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/ROMK/ROMK_pa_c3.rds")
```

```{r}
plot_intensity_distribution(romk_col, channels = c("C1.C3.hole", "C2.C3.hole"), plot_type="density", channel_colors = c(C1.C3.hole = "forestgreen", C2.C3.hole = "lightcoral"))
plot_intensity_distribution(romk_col, channels = c("C1.C3.hole", "C2.C3.hole"), plot_type="ecdf", channel_colors = c(C1.C3.hole = "forestgreen", C2.C3.hole = "lightcoral"))

ROMK_col_c3 <- assign_cell_FINAL(
  SE = romk_col,
  channels = c("C1.C3.hole", "C2.C3.hole"),
  value_col = "Mean",
  classification_method = "clustering",
  color_output = "Color_cluster",
  plot_umap = TRUE,
  k_centers = 3
)

ROMK_col_c3 <- assign_cell_FINAL(
  SE = ROMK_col_c3,
  channels = c("C1.C3.hole", "C2.C3.hole"),
  chan2analyze = c("C1.C3.hole", "C2.C3.hole"),
  value_col = "Mean",
  classification_method = "manual",
  color_output = "Color_manual",
  threshold_mode = "manual",
  green_thresh = 5.75,
  red_thresh = 4.5
)

plot_well_assignment(ROMK_col_c3, color_col = "Color_cluster")
plot_well_assignment(ROMK_col_c3, color_col = "Color_cluster", colors = c("1" = "red", "2" = "grey", "3" = "green"))
plot_well_assignment(ROMK_col_c3, color_col = "Color_manual", colors = c("red"="red", "green" = "green", "black"="grey"))

saveRDS(ROMK_col_c3, "/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/ROMK/ROMK_col_c3.rds")

```


```{r}
image_file <- "/Users/athenaschumacher/Downloads/E-Phac:Robinson_Labs/ROMK/Output_n1_ROMK_2ndRUN_28.05_22W.db"

df_img_coloc <- prepareImgDF(image_file, scale_num = TRUE, analysis = "coloc")
df_img_coloc <- df_cleaned(df_img_coloc)


# Only keep relevant ROIs
df_img_coloc <- df_img_coloc %>%
  dplyr::filter(CorrSel %in% c("Hole_ROI", "background_ROI"),
                Image_Type == "fluor")

df_coloc_background <- df_img_coloc %>%
  filter(CorrSel == "background_ROI", Image_Type == "fluor") %>%
  group_by(Plate_ID, Well, Channel_Name) %>%
  ag(cols = c("Plate_ID", "Well", "Channel_Name", "Image_Type", "CorrSel", "Second_Channel"), fun = mean)

# Sanity check
table(df_img_coloc$CorrSel)

df_img_pa <- prepareImgDF(image_file, scale_num = TRUE, analysis = "pa")
df_img_pa <- df_cleaned(df_img_pa)

df_img_pa <- df_img_pa %>%
  dplyr::filter(CorrSel %in% c("Hole_ROI", "background_ROI"),
                Image_Type == "fluor")

df_pa_background <- df_img_pa %>%
  dplyr::filter(Image_Type == "fluor", CorrSel == "background_ROI") %>%
  ag(cols = c("Date", "Plate_ID", "Well", "Channel_Name", "Image_Type", "CorrSel"), fun = mean)

table(df_img_pa$CorrSel)

romk_col <- mergeSEandImg(romk_se, df_img_coloc, tableType = "coloc",
                         Selection = "Hole_ROI", suffix = "hole")

romk_col <- mergeSEandImg(romk_col, df_coloc_background, tableType = "coloc",
                         Selection = "background_ROI", suffix = "background")

romk_pa <- mergeSEandImg(romk_se, df_img_pa, tableType = "pa",
                         Selection = "Hole_ROI", suffix = "hole")

romk_pa <- mergeSEandImg(romk_pa, df_pa_background, tableType = "pa",
                         Selection = "background_ROI", suffix = "background")


unique(romk_se$C3.C2$CorrSel)
names(colData(romk_pa))
```


Analyze the background
```{r}
extracted_db <- extractDB(image_file)
df_pa <- as.data.frame(extracted_db$pa)

df_coloc <- as.data.frame(extracted_db$coloc)

df_allSel <- df_pa %>% 
  filter(Selection == "allSelected")

df_allSel <- df_allSel %>%
  filter(Area >= 50)

C1_particles <- df_allSel %>%
  filter(Channel_Name == "C1")
median(C1_particles$Area)

C2_particles <- df_allSel %>%
  filter(Channel_Name == "C2")
median(C2_particles$Area)

C3_particles <- df_allSel %>%
  filter(Channel_Name == "C3")
median(C3_particles$Area)

ggplot(df_allSel, aes(x=Area, color = Channel_Name)) + stat_ecdf() + geom_vline(xintercept = median(C3_particles$Area), color="blue") +  geom_vline(xintercept = median(C2_particles$Area), color="green") + geom_vline(xintercept = median(C1_particles$Area), color="red")

df_allSel$NormalizedParticles <- 0

df_allSel <- df_allSel %>%
  mutate(NormalizedParticles = case_when(
    Channel_Name == "C1" ~ Area/median(C1_particles$Area),
    Channel_Name == "C2" ~ Area/median(C2_particles$Area),
    Channel_Name == "C3" ~ Area/median(C3_particles$Area),
    TRUE ~ NA_real_
  ))

df_coloc_C2 <- df_coloc %>%
  filter(Channel_Name == "C2")

df_coloc_C3 <- df_coloc %>%
  filter(Channel_Name == "C3")

unique(df_allSel$Channel_Name)
unique(df_coloc$Channel_Name)

df2 <- merge(df_allSel, df_coloc, by= c("Plate_ID", "Well", "Channel_Name", "Image_ID", "Selection"), suffix = c(".pa", ".coloc"))

Median.C1 <- df2 %>%
  filter(Channel_Name == "C1") %>%
  pull(Area.coloc) %>%
  median(na.rm = TRUE)

Median.C2 <- df2 %>%
  filter(Channel_Name == "C2") %>%
  pull(Area.coloc) %>%
  median(na.rm = TRUE)

Median.C3 <- df2 %>%
  filter(Channel_Name == "C3") %>%
  pull(Area.coloc) %>%
  median(na.rm = TRUE)

df2<- df2 %>% 
  group_by(Channel_Name)  %>%
  mutate(AreaNorm = case_when(
    Second_Channel == "C1" ~ Area.coloc/Median.C1,
    Second_Channel == "C2" ~ Area.coloc/Median.C2,
    Second_Channel == "C3" ~ Area.coloc/Median.C3
  ))

df2$NormPa <- df2$Number_of_Particles.pa / df2$Number_of_Particles.coloc

ggplot(df2, aes(x=Channel_Name, y=NormPa, fill=Second_Channel)) + geom_boxplot()


p1 <- ggplot(df, aes(x=Mean, y=Mean_C2)) +
  geom_point(alpha=0.6) + 
  geom_smooth(method = "lm", se = FALSE, color = "lightblue") +
  stat_cor(method = "pearson") + 
  labs(title = "Mean vs Mean_C2", x = "Mean", y = "Mean_C2") +
  theme_minimal()

p2 <- ggplot(df, aes(x=Mean_C2, y=Mean_C3)) +
  geom_point(alpha=0.6) + 
  geom_smooth(method = "lm", se = FALSE, color = "darkgreen") +
  stat_cor(method = "pearson") + 
  labs(title = "Mean_C2 vs Mean_C3", x = "Mean_C2", y = "Mean_C3") +
  theme_minimal()

p3 <- ggplot(df, aes(x=Mean, y=Mean_C3)) +
  geom_point(alpha=0.6) + 
  geom_smooth(method = "lm", se = FALSE, color = "darkred") +
  stat_cor(method = "pearson") + 
  labs(title = "Mean vs Mean_C3", x = "Mean", y = "Mean_C3") +
  theme_minimal()

ggpubr::ggarrange(p1, p2, p3, ncol = 3)

```



